# Topic 8 Exercises: Tree-based models

## Márcio Porto

## Programming

### 8.4.12

```{r}
set.seed(1)
library(ISLR)
data(Weekly)

train = sample(nrow(Weekly), nrow(Weekly) / 2)
# Convert qualitative to quantitative
Weekly$Direction = ifelse(Weekly$Direction == "Up", 1, 0)

Weekly.train = Weekly[train,]
Weekly.test = Weekly[-train,]

# Logistic regression
logit = glm(Direction ~ . - Year - Today, data = Weekly.train, family = "binomial")
logit_probs = predict(logit, newdata=Weekly.test, type="response")
logit_pred = ifelse(logit_probs > 0.5, 1, 0)
table(Weekly.test$Direction, logit_pred)

# Boosting
library(gbm)
boost = gbm(Direction ~ . - Year - Today, data=Weekly.train, distribution="bernoulli", n.trees=5000)
boost_probs = predict(boost, newdata=Weekly.test, n.trees=5000)
boost_pred = ifelse(boost_probs > 0.5, 1, 0)
table(Weekly.test$Direction, boost_pred)

# Bagging
library(randomForest)
bag.fit <- randomForest(Direction ~ . - Year - Today, data = Weekly.train, mtry = 6)
bag.probs <- predict(bag.fit, newdata = Weekly.test)
bag.pred <- ifelse(bag.probs > 0.5, 1, 0)
table(Weekly.test$Direction, bag.pred)

# Random forests
rf.fit <- randomForest(Direction ~ . - Year - Today, data = Weekly.train, mtry = 2)
rf.probs <- predict(rf.fit, newdata = Weekly.test)
rf.pred <- ifelse(rf.probs > 0.5, 1, 0)
table(Weekly.test$Direction, rf.pred)
```

LR classification error = 0.4623853
Boosting classification error = 0.4954128
Bagging classification error = 0.7137615
Random forests classification error = 0.4550459

Random forests is the best model for this data since it give the lowest classification error. Linear regression was not too far behind.

## Theory

### 8.4.1

```{r}

```

### 8.4.2

In this case, each term involves only a single variable, which is exactly what an additive model does. In an additive model, each new step adds more split to the tree, which is the same for boosting when d=1.

### 8.4.3

```{r}
p = seq(0, 1, 0.01)
gini = 2 * p * (1 - p)
classification_error = 1 - pmax(p, 1 - p)
cross_entropy = - (p * log(p) + (1 - p) * log(1 - p))

plot(p, classification_error, type = "l", ylim = c(0, 0.75))
lines(p, gini, col = "blue")
lines(p, cross_entropy, col = "green")
```

### 8.4.4

#### a)

If X1 >= 1 then 5. <br>
Else if X2 >= 1 then 15. <br>
Else if X1 < 0 then 3. <br>
Else if X2 < 0 then 10. <br>
Else 0.

#### b)

```{r}
par(xpd=NA)

plot(NA, NA, type="n", xlim=c(-2, 2), ylim=c(-3, 3), xlab="X1", ylab="X2")

lines(x=c(-2, 2), y=c(1, 1))

lines(x=c(1, 1), y=c(-3, 1))
text(x=(-2 + 1)/2, y=-1, labels=c(-1.8))
text(x=1.5, y=-1, labels=c(0.63))

lines(x=c(-2, 2), y=c(2, 2))
text(x=0, y=2.5, labels=c(2.49))

lines(x=c(0, 0), y=c(1, 2))
text(x=-1, y=1.5, labels=c(-1.06))
text(x=1, y=1.5, labels=c(0.21))
```

### 8.4.5

Majority vote: X = red, since most of the observations are red. <br>
Average probability: X = green, because the average of all observations is 0.45.




