# Topic 4 Exercises: Classification

## Márcio Porto

### Programming Assignment

#### 4.7.11

##### a)

```{r}
library(class)
library(ISLR)
library(MASS)

attach(Auto)
mpg01 = rep(0, length(mpg))
mpg01[mpg > median(mpg)] = 1
Auto = data.frame(Auto, mpg01)
```

##### b)

```{r}
boxplot(cylinders ~ mpg01, data=Auto, main="Cylinders and mpg01")
boxplot(displacement ~ mpg01, data=Auto, main="Displacement and mpg01")
boxplot(horsepower ~ mpg01, data=Auto, main="Horsepower and mpg01")
boxplot(weight ~ mpg01, data=Auto, main="Weight and mpg01")
boxplot(acceleration ~ mpg01, data=Auto, main="Acceleration and mpg01")
boxplot(year ~ mpg01, data=Auto, main="Year and mpg01")
```

Cylinders, weight, displacement and horsepower might help predict mpg01.

##### c)

```{r}
train = (year %% 2 == 0)
Auto.train = Auto[train,]
Auto.test = Auto[!train,]
mpg01.test = mpg01[!train]
```

##### d)

```{r}
lda = lda(mpg01 ~ cylinders + weight + displacement + horsepower, data=Auto, subset=train)
summary(lda)

pred = predict(lda, Auto.test)

table(pred$class, mpg01.test)
mean(pred$class != mpg01.test)
```

Test error rate = 12.63736%

##### e)

```{r}
qda = qda(mpg01 ~ cylinders + weight + displacement + horsepower, data=Auto, subset=train)
summary(qda)

pred_qda = predict(qda, Auto.test)
table(pred_qda$class, mpg01.test)

mean(pred_qda$class != mpg01.test)
```

Test error rate = 13.18681%

##### f)

```{r}
glm = glm(mpg01 ~ cylinders + weight + displacement + horsepower, data=Auto, family=binomial, subset=train)
summary(glm)

probs = predict(glm, Auto.test, type = "response")
pred_glm = rep(0, length(probs))
pred_glm[probs > 0.5] = 1
table(pred_glm, mpg01.test)

mean(pred_glm != mpg01.test)
```

Test error rate = 12.08791%

##### g)

```{r}
train_X = cbind(cylinders, weight, displacement, horsepower)[train,]
test_X = cbind(cylinders, weight, displacement, horsepower)[!train,]
train_mpg01 = mpg01[train]
set.seed(1)

pred_knn = knn(train_X, test_X, train_mpg01, k=1)
table(pred_knn, mpg01.test)
mean(pred_knn != mpg01.test)

pred_knn = knn(train_X, test_X, train_mpg01, k=10)
table(pred_knn, mpg01.test)
mean(pred_knn != mpg01.test)

pred_knn = knn(train_X, test_X, train_mpg01, k=100)
table(pred_knn, mpg01.test)
mean(pred_knn != mpg01.test)
```

K=1, test error rate = 15.38462%
K=10, test error rate = 16.48352%
K=100, test error rate = 14.28571%

#### 4.7.13

```{r}
attach(Boston)

crim01 = rep(0, length(crim))
crim01[crim > median(crim)] = 1
Boston = data.frame(Boston, crim01)

train = 1:(length(crim) / 2)
test = (length(crim) / 2 + 1):length(crim)
Boston.train = Boston[train,]
Boston.test = Boston[test,]
crim01.test = crim01[test]

# GLM
glm = glm(crim01 ~ . - crim01 - crim, data=Boston, family=binomial, subset=train)
probs = predict(glm, Boston.test, type="response")
pred_glm = rep(0, length(probs))
pred_glm[probs > 0.5] = 1
table(pred_glm, crim01.test)
mean(pred_glm != crim01.test)

# GLM 2
glm2 = glm(crim01 ~ . - crim01 - crim - chas - nox, data = Boston, family = binomial, subset = train)
probs = predict(glm2, Boston.test, type="response")
pred_glm = rep(0, length(probs))
pred_glm[probs > 0.5] = 1
table(pred_glm, crim01.test)
mean(pred_glm != crim01.test)

# LDA
lda = lda(crim01 ~ . - crim01 - crim, data=Boston, subset=train)
pred_lda <- predict(lda, Boston.test)
table(pred_lda$class, crim01.test)
mean(pred_lda$class != crim01.test)

#KNN
train_X = cbind(zn, indus, chas, nox, rm, age, dis, rad, tax, ptratio, black, lstat, medv)[train,]
test_X = cbind(zn, indus, chas, nox, rm, age, dis, rad, tax, ptratio, black, lstat, medv)[test,]
train_crim01 = crim01[train]

set.seed(1)

pred_knn = knn(train_X, test_X, train_crim01, k=1)
table(pred_knn, crim01.test)
mean(pred_knn != crim01.test)

pred_knn = knn(train_X, test_X, train_crim01, k=10)
table(pred_knn, crim01.test)
mean(pred_knn != crim01.test)

pred_knn = knn(train_X, test_X, train_crim01, k=100)
table(pred_knn, crim01.test)
mean(pred_knn != crim01.test)
```

GLM test error rate = 18.18182% <br>
GLM2 test error rate = 15.81028% <br>
LDA test error rate = 13.43874% <br>
K=1 KNN test error rate = 45.8498% <br>
K=10 KNN test error rate = 11.85771% <br>
K=100 KNN test error rate = 49.01186% <br>

The model with the lowest error rate is a KNN with k=10.

### Theory Assignment

#### 4.7.1

p(X) = (e^(B0 + B1X)) / (1 + e^(B0 + B1X))
p(X) = (e^(B0 + B1X)) * (1 - p(X))
p(X) / (1 - p(X)) = e^(B0 + B1X)

#### 4.7.8

It is better to choose logistic regression. The average error rate of 18% hides the fact that the training error rate is 0% for KNN (which in turn make the test error rate 36%). That being the case, the test error rate for KNN is higher than that for logistic regression, so we should choose the latter.

#### 4.7.9

##### a)

```{r}
px = 0.37 / (1 + 0.37)
px
```

##### b)

```{r}
px = 0.16
px / (1 - px)
```

